Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-06-12 Manuscript/proto-manuscript.tex
Updates

The scaling of these partial charges by a factor of 0.8 reduce the 
area per molecule to ??, which is significantly smaller than the
experimental and original Lipid 14 values.  
The decrease of area was speculated 
\todo{Did you analyze the effect of this to hydration or did we only speculate?}
to arise from reduced hydration
of headgroup due to the lower polarity of molecules with scaled charges.
The hydration can be increased by decreasing the radius of atoms
by reducing $\sigma$ term in Lennart-Jones potential.
This allows water molecules approach closer to lipid atoms,
where electrostatic interactions are stronger and, thus, 
increases hydration level.
\todo{This effect may have an official name. In that case we should mention it.}
\todo{Is there some justification for the below sentence?} 
In line with the electronic continnum correction,
the scaling factor for sigma parameters, $f_\sigma$,
has to lay between $f_q$ and 1. Scaling factor
value of ?? for $\sigma$ was found to reproduce
area per molecule values close to experiments and
original Lipid14 model when charges were scaled with
0.8.

@@

\subsection{Lipid bilayers without ions}
The structural quality of
new ECC lipid model is evaluated against scattering form factors and NMR order parameters
in Fig. \ref{simVSexpNOions}
The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
This ensures that the presented model describes correctly the membrane strucutre,
adopts a liquid disordered phase, 
and provides area per lipid agreeing with experiments (see Fig~\ref{simVSexpNOions} and Table~\ref{tab:apls}).

Order parameter changes from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
as a function of bound charge are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}. 
Approximately linear decrease of headgroup order parameters is observed both 
in simulations and experiments with the cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$). 
The use of a cationic surfactant has the benefit that it is not a subject of partitioning between water and lipid phase. 
This means that we know the exact amount of bound charge in the membrane both in experiment and simulation. 
Quantitative comparison reveals that in the original Lipid14 model the response is overestimated for both order parameters, $\alpha$ and $\beta$. 
This suggests that the observed overestimated response in \cite{catte16} 
is at least in part due to the high sensitivity of the headgroup order parameter response, 
not only due to overestimated binding of cations. 
Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in 
experiments (VALexp vs VALsim\todoi{Add values of $S^\alpha/S^\beta$ response for sim and experiment}).
This property is kept also by the newly derived model, ECC-lipid, and, hence, 
it shows a perfect agreement in the response of the headgroup order parameter $S^\alpha$ 
but underestimates slightly the response of $S^\beta$. 
\todo{Add acyl chain order parameters, POPC chemical structure}
\todo{Why original Lipid14 model seems to deviate from experiments, in contrast to the original publication? Different simulation setup used?}
\begin{figure}[]
  \centering
  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
%  \caption{\label{fig:form-f}
%    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
%\end{figure}
%
%\begin{figure}[]
%  \centering
  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
  \caption{\label{simVSexpNOions}
    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models. 
    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and EECLipid17 models
    compared with experimental order parameters from \cite{ferreira13}.}
\end{figure}
Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
are shown in Table \ref{tab:apls}.
\todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
\begin{table}
  \caption{Area per lipid from different models for POPC without ions\label{tab:apls} }
  \begin{tabular}{c c c}
    model          & A (Å$^2$)   & Temperature [K] \\
    \hline
    Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
    Lipid14ecc0.80+sigma0.875 &        &  313    \\
    GMX small patch           & 64.9   &         \\
    GMX 4xbig patch           & 65.5   &         \\
    oMM small patch           & 63.65  &         \\
    oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment \cite{Jambeck2012}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
    experiment  & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular}
\end{table}

@@

>>>>

The scaling of these partial charges by a factor of 0.8 reduce the 
area per molecule to ??, which is significantly smaller than the
experimental and original Lipid 14 values.  
The decrease of area was speculated 
\todo{Did you analyze the effect of this to hydration or did we only speculate?}
to arise from reduced hydration
of headgroup due to the lower polarity of molecules with scaled charges.
The hydration can be increased by decreasing the radius of atoms
by reducing $\sigma$ term in Lennart-Jones potential, which
allows water molecules approach closer to lipid atoms and
have stonger electrostatic interactions with them.
\todo{This effect may have an official name. In that case we should mention it.}
Indeed, the scaling of $\sigma$ increased the area per molecule
and the results close to experimental and original Lipid14 values were achieved
with scaling factor of ?? for $\sigma$ and 0.8 for partial charges.
%\todo{Is there some justification for the below sentence?} 
%In line with the electronic continnum correction,
%the scaling factor for sigma parameters, $f_\sigma$,
%has to lay between $f_q$ and 1.

@@

\subsection{Lipid bilayers without ions}
The scattering form factors, NMR order parameters and area per lipids calculated from 
the ECC corrected lipid model for POPC are compared to experiments and original
Lipid14 results in Fig. \ref{simVSexpNOions} and in Table~\ref{tab:apls}. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
The structural quality is comparable to the state of art lipid models available in literature \cite{ollila16},
thus we conclude that the ECC corrected lipid model reproduces the lipid bilayer structure
in liquid disordered phase with similar accuracy than other available models
\todo{Discussion to be finished when we have all the results in the figure}.

Headgroup order parameter response to bound charge was evaluated against
experimental data measured with cationic surfactant
(dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) \cite{scherer89}.
The exact amount of bound charge in the membrane is known these systems, because
practically all cationic surfactant molecules are embedded in lipid bilayer due to
their amphiphilic nature. Thus, the headgroup order parameter changes as a function of
mole fraction of cationic surfactants in  Fig. \ref{OrderParameterCHANGESnewMODELS} gives
also the order parameter changes as a function of bound cations.
\todo{I think that we should make a separate figure for this one and ion concentrations.
  In the current figure it would be difficult to show this with a reasonable scale.
  Currently Lipid14 results are cut out, thus the scale should be increased.
  This would, however, make the changes with CaCl too small.
  The change of headgroup tilt could be incorporated in the same figure with order parameter
  response to cationic surfactant.
}
The headgroup order parameter response to bound cation concentration
is approximately linear up to $\sim$0.3 mole fraction in experiments \cite{scherer89}.
The linearity is also observed in simulations with original Lipid14 and with ECC correction.
%from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
%as a function of bound charge are shown in
Quantitative comparison, however, reveals that the response is overestimated in 
original Lipid14 for both segments, while the ECC corrected model gives a
good agreement for the change of $\alpha$ segment order parameter, but slightly
underestimates the $\beta$ segment order parameter change.
The overestimation of order parameter changes with original Lipid14 model 
suggests that the overestimated response with CaCl$_2$ concentration in \cite{catte16} 
can be partly explained by the high sensitivity of the headgroup order parameter response
to bound charge.

Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in 
experiments (VALexp vs VALsim
\todo{Add values of $S^\alpha/S^\beta$ response for sim and experiment
  SAMULI: Maybe these could be put in the same figure with the order parameter
  response and P-N vector angle change with surfactants.
}
).
This property is kept also by the newly derived model, ECC-lipid, and, hence, 
it shows a perfect agreement in the response of the headgroup order parameter $S^\alpha$ 
but underestimates slightly the response of $S^\beta$. 
\begin{figure}[]
  \centering
  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
%  \caption{\label{fig:form-f}
%    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
%\end{figure}
%
%\begin{figure}[]
%  \centering
  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
  \caption{\label{simVSexpNOions}
    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models. 
    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and EECLipid17 models
    compared with experimental order parameters from \cite{ferreira13}.}
  \todo{Add acyl chain order parameters, POPC chemical structure}
  \todo{Should we add the results from original lipid14?}
\end{figure}
Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
are shown in Table \ref{tab:apls}.
\todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
\begin{table}
  \caption{Area per lipid from different models for POPC without ions\label{tab:apls} }
  \begin{tabular}{c c c}
    model          & A (Å$^2$)   & Temperature [K] \\
    \hline
    Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
    Lipid14ecc0.80+sigma0.875 &        &  313    \\
    GMX small patch           & 64.9   &         \\
    GMX 4xbig patch           & 65.5   &         \\
    oMM small patch           & 63.65  &         \\
    oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment \cite{Jambeck2012}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
    experiment  & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular}
\end{table}

@@

not only due to overestimated binding of cations.
